---
title: "Plots and Figures for Comparative Hummingbird Blood paper"
author: "Jessie Williamson"
date: "7/21/2020"
output:
  md_document:
    variant: markdown_github
---

#### 

PLOTS & FIGS SCRIPT 

####

This script contains exploratory plots, polished plots, and figures for allhummingbird_blood_comparison.Rmd. These are separated into a standalone script to not bog down main script. 


# Load packages
```{R}
library(reshape)
library(reshape2)
library(plyr)
library(dplyr)
library(car)
library(GGally)
library(Hmisc)
library(gridExtra)
library(stats)
library(gplots)
library(ggplot2)
library(stats4) # Forces knitr to work when it's being wonky
library(PMCMR) #Allows Kruskal-Wallis post-hocs
library(effects)
library(gridExtra)
library(lattice)
library(survival)
library(fmsb)
library(faraway)
library(ape)

# Mapping 
library(raster)
library(sp)
library(rgdal)
library(RStoolbox)
library(prettymapr)
library(viridis)
library(rasterVis)

# Modeling packages 
library(nlme)
library(lme4)
library(AICcmodavg)
library(MuMIn)
library(glmulti)
library(lsmeans)
library(rsq) # get r-squared values from GLM
library(r2glmm) # for R^2 values from lmer() and glmer()
library(multcompView) # related to multiple comparisons?
library(jtools) # interaction plots 
library(interactions) # interaction plots 
library(broom)
library(stargazer) # model output tables
library(ggeffects) # for estimating model predictions from mixed effects models
library(brms)
library(MCMCglmm)
library(rstan)
library(bayesplot)

# Phylo packages 
library(phytools)
```


---

# Clear workspace and set WD
```{R}
rm(list=ls(all=TRUE)) # clear workspace 
setwd("/Users/Jessie/Dropbox (MSBbirds)/Rdirectory/ComparativeHummingbirdBlood")
```



# load in data 
```{R}
# # Load functions 
# source("ada_functions.R") # Erik's ADA functions for clean & collated lm diag plots

# Final hummingbird blood dataset
# Don't need to load if already loaded from another script 
final <- read.csv("AllHummingbirdBlood_Wrangled_withBioClim_andPCA_03-01-21.csv")
final <- final[ , !(colnames(final) %in% c("X"))] # drop weird X column 1 if reading in from read.csv

# Patagona-free subset (for possible modeling of mass, where Patagona is a huge body size outlier):
final.nopgig <- read.csv("HumBlood.final.nopgig_03-01-21.csv")
final.nopgig <- final.nopgig[ , !(colnames(final.nopgig) %in% c("X"))] # drop weird X column 1 if reading in from read.csv

# Phylogeny pruned from McGuire and adjusted (see script HumBlood_Phylogeny.Rmd)
#tree <- read.tree("McGuirePruned_AllHummingbirdComparativeTree_FINAL.tre") # a list 
# tree$tip.label # Check 77 tips
# plotTree(tree, ftype="i") # Check this pre-final tree w/ adjusted branch lengths 
```


# Set levels to ensure that colors plot properly
Make sure levels correspond to clade label order in phylogeny from top to bottom
```{r}
# Set clade levels/make sure these are correct
final$Clade <- factor(final$Clade
                       , levels = c("Coquette",
                                    "Brilliant", 
                                    "Patagona", 
                                    "Bee", 
                                    "Mountain-gem", 
                                    "Emerald", 
                                    "Mango",  
                                    "Topaz",
                                    "Hermit"))
```



# COLOR PALETTES 
```{r}
# Colors from Figure 1 map phylo fig (last revised 2/23/21); see Illustrator file
# Turquoise - 4FBBB6 (coquette) **may need to change because very similar to points on the map
# Gray - A5A5AC (brilliant)
# Red - EE4024 (patagona)
# Gold - D9CA44 (bees)
# Pink - E786B7 (mountain gems)
# Green - 37B453 (emeralds)
# Orange - F9AC1B (mangos)
# Blue - 00A5CB (topazes)
# purple - 9E6EAE (hermits)

# HUMMINGBIRD COLORS 
# hum_colors <- c( "#4FBBB6", # Turquoise (Coquette)
#                  "#A5A5AC", # gray (Brilliant)
#                  "#EE4024", # red (Patagona)
#                  "#D9CA44", # Gold (Bee)
#                  "#E786B7", # pink (Mountain-gem)
#                  "#37B453", # green (Emerald)  
#                  "#F9AC1B", # orange (Mango)
#                  "#00A5CB", # blue (Topaz)
#                  "#9E6EAE" # purple (Hermit)
#                             )

# revised 
hum_colors <- c( "#04A5A5", # Turquoise (Coquette)
                 "#9E9285", # brown (Brilliant)
                 "#EE4024", # red (Patagona)
                 "#F4CD0B", # Gold (Bee)
                 "#E786B7", # pink (Mountain-gem)
                 "#97B25B", # green (Emerald)  
                 "#F79E0F", # orange (Mango)
                 "#03446B", # navy blue (Topaz)
                 "#916EF9" # purple (Hermit)
                            )

# Oreotrochilus chimborazo color palette: 
# Deep purple: #6E1EDC
# Aqua gorget: #1CD2D8; another color: #0DAFA9
# pollen forehead: #D49A76
# gray patch brown: #665E54; another version #726A61; another version #8C847B
# deep green: #344D3F
# chuquiragua buds: #E2833F; #F4893C
# chuquiragua leaves: #6F8F17
# snowy breast: #D1D1D9
# Iridescent cheeks: #BA51FA
#scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87

# Ochimborazo_palette <- c( "#0E0E17", # Black 
#                           "#1CD2D8", # turquoise aqua gorget
#                           "#8C847B", # browny flanks 
#                           "#F4893C", # Chuquiragua buds 
#                           "#166975", # Darker part of gorget
#                           "#FDC31F", # Chuquiragua pollen  
#                           "#BA51FA", # Pinkish shine to gorget; Dark green wing patches: #344D3F
#                           "#6E1EDC", # Deep purple 
#                           "#6F8F17" # Chuquiragua leaves 
#                             )
# 
# peru_palette <- c( "#628454", # sage green
#                    "#FB4301", # deep orange-red
#                    "#6D7B9C", # skirt blue
#                    "#966D56", # Adobe brown
#                    "#E7860E", # golden orange-yellow
#                    "#264142", # Deep sea green
#                    "#BFD1F8", # periwinkle on shirt
#                    "#952842", # cabernet
#                    "#F871AA" # pink
#                             )
# 
# # Peru chimborazo: inspired by Peruvian textiles and Oreotrochilus chimborazo colors 
# peru_chimborazo <- c( "#FDC31F", # Chuquiragua pollen
#                       "#1CD2D8", # turquoise aqua gorget
#                       "#8C847B", # browny flanks 
#                       "#9AB243", # Chuquiragua gree
#                        "#FB4301", # Deep orange-red
#                       "#264142", # Deep sea green
#                       "#F871AA", # pink
#                       "#6E1EDC", # Deep purple 
#                       "#6D7B9C" # skirt blue # played around w/ subbing in this too:   "#A8A9AD" # gray breast
#                             )
```


# Basic plot of sampling across elevations
```{r}
# SUPER BASIC PLOTS, missing a lot of formatting/aesthetics 
# Sampling density across elevation 
(p <- ggplot(final, aes(x=elev, y=..density..)) +
      geom_density(alpha=.2, fill="#FF6666") + 
      geom_histogram(binwidth = 200) + 
    #  scale_color_manual(values = hum_colors) +
      theme_classic() + 
      labs(x="Elevation (m)", y="Sampling Density") + 
      theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )

# Sampling Depth across elevation  
(p <- ggplot(final, aes(x=elev)) +
   #   geom_density(alpha=.2, fill="#FF6666") + 
      geom_histogram(binwidth = 200) + 
    #  scale_color_manual(values = hum_colors) +
      theme_classic() + 
      labs(x="Elevation (m)", y="Sampling Depth") + 
      theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )

# Overlaid histograms
# ggplot(final, aes(x=elev)) +
#   geom_histogram(fill="black", alpha=0.5, position="identity")
# 
# ggplot(final, aes(elev, colour = Clade)) +
#   geom_freqpoly(binwidth = 500)
# 
# ggplot(final, aes(elev)) +
#   geom_histogram(bins = 100) + 
#   theme_classic() 
```



# PLOTS OF BLOOD CHARACTERISTICS WITH ELEVATION 
Figure 2 - 6-panel fig w/ relationship between predictors and elevation 
```{r}
# Scatterplots of each of the 6 blood characteristics with elevation, colored by clade

# Code to set shapes by clade, but this was totally overwhelming
  # scale_shape_manual(values=c("Bee"=4, "Brilliant"=17, "Coquette"=15, # Shapes are a little overwhelming, honestly
  #                             "Emerald"=18, "Hermit"=25, "Mango"=8, "Mountain-gem"=3,
  #                             "Patagona"=16, "Topaz"=9)) +
 #   scale_shape_manual(values=shapes) + # Another way to do this is set colors above 

# Scatterplot: Hb v. elev 
(p1 <- ggplot(final, aes(x=elev, y=hb, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
 # geom_jitter(width=0.5, height=0.5, alpha=0.8, size=1.1) + # jitter to fill plot space 
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x=" ", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="[Hb] g/dl") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") + 
  ggtitle("A") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.2,0.1,0.0,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p1, filename="p1_ScatterplotBasic_Elev_by_Hb.pdf", height=7, width=9, units="in")
# The 166 rows of missing values come from many NA Hb values; don't worry about these 

# Scatterplot: Hct v. elev
(p2 <- ggplot(final, aes(x=elev, y=hct, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  # scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="Hct (%)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("B") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.2,0.2,0.0,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p2, filename="p3_ScatterplotBasic_Elev_by_Hct.pdf", height=7, width=9, units="in")

# Scatterplot: TRBC v. elev
(p3 <- ggplot(final, aes(x=elev, y=trbc, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
  #geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
   labs(y=expression(TRBC~"("~RBC~x~10^{6}~"/"~mm^{3}~")"), # horrible expression language for getting formatted y axis 
     x="") + # Removed "Elevation (m)" because we don't want Panel A to have label
    # in expression world, ~ is space and mathematical symbols like parenthesis need to be set of w/ " " 
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("C") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.1,0.0,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p3, filename="p3_ScatterplotBasic_Elev_by_TRBC.pdf", height=7, width=9, units="in")

# Scatterplot: mcv v. elev
(p4 <- ggplot(final, aes(x=elev, y=mcv, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation (m)", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="MCV (fl)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("D") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  #theme(legend.box.spacing = unit(c(0.2,5,0.2,0.2), "cm")) +  # top, right, bottom, left
  theme(plot.margin = unit(c(0.0,0.2,0.0,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p4, filename="p4_ScatterplotBasic_Elev_by_MCV.pdf", height=7, width=9, units="in")

# ANNOYING: space between plot area and legend is supposedly adjusted w/ "legend.box.spacing" arg but little documentation and it's
# being stubborn. I manually exported legend, will make 6-panel fig w/out legend and will add legend on in Illustrator

# Scatterplot: MCH v. elev
(p5 <- ggplot(final, aes(x=elev, y=mch, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
 # geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation (m)", y="MCH (pg)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("E") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.1,0.2,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p5, filename="p5_ScatterplotBasic_Elev_by_MCH.pdf", height=7, width=9, units="in")


# Scatterplot: MCHC v. elev
(p6 <- ggplot(final, aes(x=elev, y=mchc, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
  #geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation (m)", y="MCHC (g/dl)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("F") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.2,0.2,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p6, filename="p6_ScatterplotBasic_Elev_by_MCHC.pdf", height=7, width=9, units="in")


library(patchwork)
# Combine all plots to make one 6-panel figure:
# Fig2_BloodVElev_6panel <- ((p1 + p2) / (p3 + p4) / (p5 + p6) +  # the / literally means put p14 on top of p15
#                                   plot_layout(guides = "collect")) # "collates" and combines one common legend 
#                                   # In order for plot_layout to "collect" legend, must have printed legend for indiv plots above
# plot(Fig2_BloodVElev_6panel)
# ggsave(Fig2_BloodVElev_6panel, filename = "Fig2_BloodVElev_6panel_03-01-21.pdf", bg="transparent", height=7, width=7, units="in")
# This looks ok, but horizontal format (below) is better

# Alternative layout option = BETTER, USE THIS ONE: 
Fig2_BloodVElev_6panel_v2 <- (p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(guides = "collect"))
plot(Fig2_BloodVElev_6panel_v2)
ggsave(Fig2_BloodVElev_6panel_v2, filename = "Fig2_BloodVElev_6panel_v2_03-01-21.pdf", bg="transparent", height=7, width=11, units="in")
# Why I like this format: primary blood indices align clearly in Panels A-C; secondary indices on bottom in D-F

# Fig2_BloodVElev_6panel <- grid.arrange(p1,p2,p3,p4,p5,p6, ncol=2) # Old method 
```

Use patchwork to arrange plots: https://gotellilab.github.io/GotelliLabMeetingHacks/NickGotelli/ggplotPatchwork.html
Many say this is a better alternative to grid.arrange() 

Tutorial to using patchwork to make plots: https://patchwork.data-imaginist.com/articles/guides/layout.html


# PLOTS FOR RELATIONSHIPS AMONG BLOOD PARAMETERS
Quick workup of predictors: 
```{r}
# full dataset
p <- ggpairs(subset(final, select = c(elev, mass, hb, hct, trbc, mcv, mch, mchc))) 
print(p)
```

Mock up of Figure 3, relationships among blood parameters 
See Hawkey et al. 1991 for inspiration.
```{r}
library(viridis)

# Panel A: RBC and MCV - redo of plot in Hawkey et al. 1991
(p12 <- ggplot(final, aes(x=mcv, y=trbc)) +
    #  scale_colour_magma(discrete=FALSE) +
# p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(size=2, alpha=0.8, colour="darkred") +
    # scale_color_manual(values=hum_colors) +
    # stat_smooth(data=final, aes(y=trbc), size=1.2, method="lm") +
    # geom_jitter(width=0.8, height=0.8, alpha=0.8, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y=expression(TRBC~"("~RBC~x~10^{6}~"/"~mm^{3}~")"), # horrible expression language for getting formatted y axis 
     x="") + # Removed "Elevation (m)" because we don't want Panel A to have label
    #scale_color_manual(values=c(neither="black", generalist="burlywood2", specialist="#853FCD")) + #336B87
    #scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
    ggtitle("A") + # Assign panel number/header; this will be (a) because first in series of 3
    theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(plot.margin = unit(c(0.2,0.2,0.0,0.2), "cm")) +  # top, right, bottom, left
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p12, filename="p12_RBC_vs_MCV_Scatterplot.pdf", height=7, width=9, units="in") 

# Wow, this is amazing. See Hawkey et al. 1991 for some dynamite interpretation
# Essentially this means: species have either evolved many small erythrocytes or few large erythrocytes 
# Seems to be a sweet spot of a moderate number of a moderate sized erythrocyte


# Panel B: MCH and MCV - redo of plot in Hawksey et al. 1991
(p13 <- ggplot(final, aes(x=mcv, y=mch)) +
# p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(size=2, alpha=0.8, colour="darkred") +
    # stat_smooth(data=full1, aes(y=trbc), size=1.2, method="lm") +
    # geom_jitter(width=0.8, height=0.8, alpha=0.8, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y="MCH (pg)", 
         x="MCV (fl)") + # fix RBC title/notation
    #scale_color_manual(values=c(neither="black", generalist="burlywood2", specialist="#853FCD")) + #336B87
    #scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
    ggtitle("B") + # Assign panel number/header; this will be (a) because first in series of 3
    theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(plot.margin = unit(c(0.0,0.2,0.2,0.2), "cm")) +  # top, right, bottom, left
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p13, filename="p13_RBC_vs_MCH_Scatterplot.pdf", height=7, width=9, units="in") 

library(patchwork)

# Combine plots to make one 2-panel figure:
Fig3_BloodComparison_2panel <- (p12 / p13 +  # the / literally means put p14 on top of p15
                                  plot_layout(guides = "collect")) # "collates" and combines one common legend 
                                  # In order for plot_layout to "collect" legend, must have printed legend for indiv plots above
plot(Fig3_BloodComparison_2panel)
# NOTE: THERE IS NO LEGEND IN THIS FIGURE, SO NO LEGEND APPEARS WHEN YOU PLOT_LAYOUT "COLLECT"

ggsave(Fig3_BloodComparison_2panel, filename = "Fig3_BloodComparison_2panel_03-01-21.pdf", bg="transparent", height=6, width=5, units="in")

# Old method (but I was having trouble getting common legend
# Fig3_BloodComparison_2panel <- grid.arrange(p12,
#                                            p13,
#                                            ncol=1))
```


Fig 3 but colored by elevation 
```{R}
# Panel A: RBC and MCV - redo of plot in Hawkey et al. 1991
(p14 <- ggplot(final, aes(x=mcv, y=trbc, colour=elev)) +
      scale_colour_viridis(option="magma", discrete=FALSE) +
     # p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(size=2, alpha=0.8) +
    # scale_color_manual(values=hum_colors) +
    # stat_smooth(data=final, aes(y=trbc), size=1.2, method="lm") +
    #geom_jitter(width=0.8, height=0.8, alpha=0.8, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y=expression(TRBC~"("~RBC~x~10^{6}~"/"~mm^{3}~")"), # horrible expression language for getting formatted y axis 
     x="") + # Removed "Elevation (m)" because we don't want Panel A to have label
    #scale_color_manual(values=c(neither="black", generalist="burlywood2", specialist="#853FCD")) + #336B87
    #scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
    ggtitle("A") + # Assign panel number/header; this will be (a) because first in series of 3
    theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(plot.margin = unit(c(0.2,0.2,0.0,0.2), "cm")) +  # top, right, bottom, left
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p14, filename="p14_RBC_vs_MCV_Scatterplot_ColoredByElev.pdf", height=7, width=9, units="in") 

# Wow, this is amazing. See Hawkey et al. 1991 for some dynamite interpretation
# Essentially this means: species have either evolved many small erythrocytes or few large erythrocytes 
# Seems to be a sweet spot of a moderate number of a moderate sized erythrocyte


# Panel B: MCH and MCV - redo of plot in Hawksey et al. 1991
(p15 <- ggplot(final, aes(x=mcv, y=mch, colour=elev)) +
    scale_colour_viridis(option="magma", discrete=FALSE) +
    # p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(size=2, alpha=0.8) +
    # stat_smooth(data=full1, aes(y=trbc), size=1.2, method="lm") +
    # geom_jitter(width=0.8, height=0.8, alpha=0.8, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y="MCH (pg)", 
         x="MCV (fl)") + # fix RBC title/notation
    #scale_color_manual(values=c(neither="black", generalist="burlywood2", specialist="#853FCD")) + #336B87
    #scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
    ggtitle("B") + # Assign panel number/header; this will be (a) because first in series of 3
    theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(plot.margin = unit(c(0.0,0.2,0.2,0.2), "cm")) +  # top, right, bottom, left
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p15, filename="p15_RBC_vs_MCH_Scatterplot_ColoredByElev.pdf", height=7, width=9, units="in") 

library(patchwork)

# Combine plots to make one 2-panel figure:
Fig3_BloodComparison_2panel_v2 <- (p14 / p15 +  # the / literally means put p14 on top of p15
                                  plot_layout(guides = "collect")) # "collates" and combines one common legend 
                                  # In order for plot_layout to "collect" legend, must have printed legend for indiv plots above
plot(Fig3_BloodComparison_2panel_v2)

ggsave(Fig3_BloodComparison_2panel_v2, filename = "Fig3_BloodComparison_2panel_ColoredByElev_03-01-21.pdf", bg="transparent", height=6, width=5, units="in")

# Old method for combining plots, but I was having trouble getting a common legend
# Fig3_BloodComparison_2panel_v2 <- grid.arrange(p14,
#                                                p15,
#                                                ncol=1)

```


# PLOTS OF BLOOD CHARACTERISTICS WITH ELEVATION_POSITION 
****THIS IS FIG. 2 redone w/ elev position! 
```{r}
# Scatterplot: Hb v. elev 
(p16 <- ggplot(final, aes(x=elev_position, y=hb, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
 # geom_jitter(width=0.5, height=0.5, alpha=0.8, size=1.1) + # jitter to fill plot space 
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x=" ", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="[Hb] g/dl") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") + 
  ggtitle("A") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.2,0.1,0.0,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p16, filename="p16_ScatterplotBasic_ElevPos_by_Hb.pdf", height=7, width=9, units="in")
# The 166 rows of missing values come from many NA Hb values; don't worry about these 

# Scatterplot: Hct v. elev
(p17 <- ggplot(final, aes(x=elev_position, y=hct, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  # scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="Hct (%)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("B") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.2,0.2,0.0,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p17, filename="p17_ScatterplotBasic_ElevPosition_by_Hct.pdf", height=7, width=9, units="in")

# Scatterplot: TRBC v. elev
(p18 <- ggplot(final, aes(x=elev_position, y=trbc, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
  #geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
   labs(y=expression(TRBC~"("~RBC~x~10^{6}~"/"~mm^{3}~")"), # horrible expression language for getting formatted y axis 
     x="") + # Removed "Elevation (m)" because we don't want Panel A to have label
    # in expression world, ~ is space and mathematical symbols like parenthesis need to be set of w/ " " 
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("C") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.1,0.0,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p18, filename="p18_ScatterplotBasic_ElevPosition_by_TRBC.pdf", height=7, width=9, units="in")

# Scatterplot: mcv v. elev
(p19 <- ggplot(final, aes(x=elev_position, y=mcv, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
#  geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.8) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation Position", # Removed "Elevation (m)" because we don't want Panel A to have label
       y="MCV (fl)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("D") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  #theme(legend.box.spacing = unit(c(0.2,5,0.2,0.2), "cm")) +  # top, right, bottom, left
  theme(plot.margin = unit(c(0.0,0.2,0.0,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p19, filename="p19_ScatterplotBasic_ElevPosition_by_MCV.pdf", height=7, width=9, units="in")

# Scatterplot: MCH v. elev_pos
(p20 <- ggplot(final, aes(x=elev_position, y=mch, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
 # geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation Position", y="MCH (pg)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("E") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.1,0.2,0.2), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p20, filename="p20_ScatterplotBasic_ElevPosition_by_MCH.pdf", height=7, width=9, units="in")

# Scatterplot: MCHC v. elev_pos
(p21 <- ggplot(final, aes(x=elev_position, y=mchc, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final, size=1.1, alpha=0.8) +
  #scale_colour_viridis(discrete=TRUE) + 
  scale_color_manual(values = hum_colors) +
  #geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="black", size=1.0) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation Position", y="MCHC (g/dl)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  ggtitle("F") + # Assign panel number/header; this will be (a) because first in series of 3
  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
         plot.title = element_text(face="bold")) + # This makes panel header bold 
       # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(plot.margin = unit(c(0.0,0.2,0.2,0.1), "cm")) +  # top, right, bottom, left
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
 )
ggsave(p21, filename="p21_ScatterplotBasic_ElevPosition_by_MCHC.pdf", height=7, width=9, units="in")


library(patchwork)
# Combine all plots to make one 6-panel figure:
# Fig2_BloodVElev_6panel <- ((p1 + p2) / (p3 + p4) / (p5 + p6) +  # the / literally means put p14 on top of p15
#                                   plot_layout(guides = "collect")) # "collates" and combines one common legend 
#                                   # In order for plot_layout to "collect" legend, must have printed legend for indiv plots above
# plot(Fig2_BloodVElev_6panel)
# ggsave(Fig2_BloodVElev_6panel, filename = "Fig2_BloodVElev_6panel_03-01-21.pdf", bg="transparent", height=7, width=7, units="in")
# This looks ok, but horizontal format (below) is better

# Alternative layout option = BETTER, USE THIS ONE: 
Fig_BloodVElevPosition_6panel <- (p16 + p17 + p18 + p19 + p20 + p21 + plot_layout(guides = "collect"))
plot(Fig_BloodVElevPosition_6panel)
ggsave(Fig_BloodVElevPosition_6panel, filename = "Fig_BloodVElevPosition_6panel_03-01-21.pdf", bg="transparent", height=7, width=11, units="in")
# Why I like this format: primary blood indices align clearly in Panels A-C; secondary indices on bottom in D-F
```








# Trial with species mean values 
If you want to do this, need to read in species mean value dataset from modeling rmd file 
```{r}
# Scatterplot: Hb v. elev 
# (p <- ggplot(species.means, aes(x=mean_mass, y=mean_hb, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
#   geom_point(size=1.5, alpha=0.8) +
#   #scale_colour_viridis(discrete=TRUE) + 
#   scale_color_manual(values = hum_colors) +
#  # geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # Don't jitter; creates the appearance of duplicate points
#   geom_smooth(method = "lm", colour="firebrick3", size=1.0) + # line used to be dark turquoise 
#   theme_classic() + 
#   labs(x="Mass (g)", y="[Hb] g/dl") +
#        # theme(axis.text.x = element_text(hjust = 1)) +
#   theme(legend.position = "right") + 
#   ggtitle("(a)") + # Assign panel number/header; this will be (a) because first in series of 3
#   theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
#          plot.title = element_text(face="bold")) + # This makes panel header bold 
#        # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
#   theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=12))
#  )
#ggsave(p, filename="p1_ScatterplotBasic_Elev_by_Hb.pdf", height=7, width=9, units="in")
```



# PRELIM DATA EXPLORATION 
```{r}

# Boxplot of differences in hb w/ species
(p22 <- ggplot(final, aes(x=Clade, y=hb)) +
       # facet_wrap(~sex, nrow=2) + 
        geom_boxplot(aes(fill = factor(Clade)), outlier.size=2.0, alpha=0.7) +
        geom_jitter(position = position_jitter(width = 0.01), alpha = 0.5, size = 2.0) +
        stat_summary(fun=mean, colour="firebrick1", geom="point", shape=18, size=6) + # Red diamonds at mean 
        theme_classic() +   # theme_classic() removes all gridlines; theme_bw() retains subtle nice gridlines           
        theme(panel.grid.minor = element_blank()) +
        labs(y="[Hb] g/dl", # Hella confusing formatting for x and y axis labels 
             x="Species") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        theme(legend.position = "none") +
        theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
#ggsave(p1, filename="p1_Boxplot_Hb_by_Clade.pdf", height=7, width=9, units="in")

# Boxplot of differences in hb w/ genotype
(p23 <- ggplot(final, aes(x=b13b83.genotype, y=hb)) +
       # facet_wrap(~sex, nrow=2) + 
        geom_boxplot(aes(fill = factor(b13b83.genotype)), outlier.size=2.0, alpha=0.7) +
        geom_jitter(position = position_jitter(width = 0.01), alpha = 0.5, size = 2.0) +
        stat_summary(fun=mean, colour="firebrick1", geom="point", shape=18, size=6) + # Red diamonds at mean 
        theme_classic() +   # theme_classic() removes all gridlines; theme_bw() retains subtle nice gridlines           
        theme(panel.grid.minor = element_blank()) +
        labs(y="[Hb] g/dl", # Hella confusing formatting for x and y axis labels 
             x="Species") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        theme(legend.position = "none") +
        theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
#ggsave(p1, filename="p1_Boxplot_Hb_by_Clade.pdf", height=7, width=9, units="in")
 


# Extremely basic plot of elev range summary
# Maybe interesting to get mean Hb value per species and plot range on x-axis and mean Hb on y? 
plot(elev.range.summary)

# Basic scatterplot of mass v. Hb 
(p <- ggplot(final, aes(x=mass, y=hb, colour=Clade)) + # uncenter Tamin values
  geom_point() +
  scale_colour_viridis(discrete=TRUE) +
  geom_smooth(method = "lm", colour="darkturquoise", size=1.2) + 
  theme_classic() + 
  labs(x="Mass (g)", y="[Hb] g/dl") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "right") +
  # ggtitle("(a)") + # Assign panel number/header; this will be (a) because first in series of 3
  # theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
  #       plot.title = element_text(face="bold")) + # This makes panel header bold 
  #     # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p3, filename="p3_ScatterplotBasic_Mass_by_Hb.pdf", height=7, width=9, units="in")
# This is hilarious: this plot is like "Patagona"...and everything else. 
# This brings up questions about how to analyze mass data w/ Patagona involved and whether this 


# Basic scatterplot of mass v. Hb 
(p <- ggplot(final.nopgig, aes(x=mass, y=hb)) + # uncenter Tamin values
  geom_point() +
  geom_smooth(method = "lm", colour="darkturquoise", size=1.2) + 
  theme_classic() + 
  labs(x="Mass (g)", y="[Hb] g/dl") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "none") +
  # ggtitle("(a)") + # Assign panel number/header; this will be (a) because first in series of 3
  # theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
  #       plot.title = element_text(face="bold")) + # This makes panel header bold 
  #     # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
   )
ggsave(p4, filename="p4_ScatterplotBasic_Mass_by_Hb_NOPATAGONA.pdf", height=7, width=9, units="in")
# Def interesting trend of increasing Hb w/ mass 


# Break these plots out by species (I prefer clade; ~species is too much and a shitshow to look at)
# not sure this is too useful except to show sampling across elev gradient 
(p5 <- ggplot(final, aes(x=elev, y=hb)) +
  facet_wrap(~Clade, scales="free") +
  geom_point(pch=21, stroke=1, aes(color=Clade), show.legend = FALSE) +
  geom_smooth(method="lm", se=FALSE, linetype="dashed", color="black") +
  theme_classic() +
  labs(x="Elevation (m)", y="[Hb] g/dl") +
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p5, filename="p5_Scatterplot_Elev_by_Hb_byCLADE.pdf", height=7, width=9, units="in")


# now just by po2 specialists: 
# Break these plots out by species 
# not sure this is too useful except to show sampling across elev gradient 
(p5.1 <- ggplot(po2.specialist, aes(x=elev, y=hb)) +
  facet_wrap(~species, scales="free") + # split by species 
  geom_point(pch=21, stroke=1, aes(color=clade), show.legend = FALSE) + # color by clade 
  geom_smooth(method="lm", se=FALSE, linetype="dashed", color="black") +
  theme_classic() +
  labs(x="Elevation (m)", y="[Hb] g/dl") +
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p5.1, filename="p5.1_Scatterplot_Elev_by_Hb_PO2-specialists.pdf", height=7, width=9, units="in")


# now just by po2 generalists: 
# Break these plots out by species 
# not sure this is too useful except to show sampling across elev gradient 
(p5.2 <- ggplot(po2.generalist, aes(x=elev, y=hb)) +
  facet_wrap(~species, scales="free") + # split by species 
  geom_point(pch=21, stroke=1, aes(color=clade), show.legend = FALSE) + # color by clade 
  geom_smooth(method="lm", se=FALSE, linetype="dashed", color="black") +
  theme_classic() +
  labs(x="Elevation (m)", y="[Hb] g/dl") +
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p5.2, filename="p5.2_Scatterplot_Elev_by_Hb_PO2-generalists.pdf", height=7, width=9, units="in")
# Huh, interesting for C. coruscans and R. vesper: both are generalists because found <1000 m
# but we've never actually sampled them at lowest elevs, hence both starting on x-axis here >2000 m




# Break these plots out by species but remove colors to simplify 
# not sure this is too useful except to show sampling across elev gradient 
(p6 <- ggplot(final, aes(x=elev, y=hb)) +
  facet_wrap(~Clade, scales="free") +
  geom_point(stroke=1, show.legend = FALSE) +
  geom_smooth(method="lm", se=FALSE, color="turquoise") +
  theme_classic() +
  labs(x="Elevation (m)", y="[Hb] g/dl") +
  theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p6, filename="p2_Scatterplot_Elev_by_Hb_byCLADE_simple.pdf", height=7, width=9, units="in")



# Boxplot of differences in hb w/ species
(p7 <- ggplot(final, aes(x=sex, y=hb)) +
       # facet_wrap(~sex, nrow=2) + 
        geom_boxplot(aes(fill = factor(sex)), outlier.size=2.0, alpha=0.7) +
        geom_jitter(position = position_jitter(width = 0.01), alpha = 0.5, size = 2.0) +
        stat_summary(fun=mean, colour="firebrick1", geom="point", shape=18, size=6) + # Red diamonds at mean 
        theme_classic() +   # theme_classic() removes all gridlines; theme_bw() retains subtle nice gridlines           
        theme(panel.grid.minor = element_blank()) +
        labs(y="[Hb] g/dl", # Hella confusing formatting for x and y axis labels 
             x="Sex") +
   #    labs(x="Season",y="Tbmin (°C)") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        theme(legend.position = "none") +
  #      ggtitle("(d)") + # Assign panel number/header
  #      theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
   #           plot.title = element_text(face="bold")) + # This makes panel header bold 
      # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
        theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p7, filename="p7_Boxplot_Hb_by_Sex.pdf", height=7, width=9, units="in")
# Good news is that Hb pretty similar between sexes 
# Might consider removing unknowns to plot this cleanly


# Hb differences w/ age boxplot 
(p8 <- ggplot(final, aes(x=age, y=hb)) +
       # facet_wrap(~sex, nrow=2) + 
        geom_boxplot(aes(fill = factor(age)), outlier.size=2.0, alpha=0.7) +
        geom_jitter(position = position_jitter(width = 0.01), alpha = 0.5, size = 2.0) +
        stat_summary(fun=mean, colour="firebrick1", geom="point", shape=18, size=6) + # Red diamonds at mean 
        theme_classic() +   # theme_classic() removes all gridlines; theme_bw() retains subtle nice gridlines           
        theme(panel.grid.minor = element_blank()) +
        labs(y="[Hb] g/dl", # Hella confusing formatting for x and y axis labels 
             x="Sex") +
   #    labs(x="Season",y="Tbmin (°C)") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        theme(legend.position = "none") +
  #      ggtitle("(d)") + # Assign panel number/header
  #      theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
   #           plot.title = element_text(face="bold")) + # This makes panel header bold 
      # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
        theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p8, filename="p8_Boxplot_Hb_by_Age.pdf", height=7, width=9, units="in")
# Needs massaging to clean this up
# Need to assign NAs to unknowns
# Adults a bit higher than juvs, but not sure differences are significant (they might be tho)
# Add age as a random effect in models? 


(p9 <- ggplot(final, aes(x=locality, y=hb)) +
       # facet_wrap(~sex, nrow=2) + 
        geom_boxplot(aes(fill = factor(locality)), outlier.size=2.0, alpha=0.7) +
        geom_jitter(position = position_jitter(width = 0.01), alpha = 0.5, size = 2.0) +
        stat_summary(fun=mean, colour="firebrick1", geom="point", shape=18, size=6) + # Red diamonds at mean 
        theme_classic() +   # theme_classic() removes all gridlines; theme_bw() retains subtle nice gridlines           
        theme(panel.grid.minor = element_blank()) +
        labs(y="[Hb] g/dl", # Hella confusing formatting for x and y axis labels 
             x="Locality") +
   #    labs(x="Season",y="Tbmin (°C)") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        theme(legend.position = "none") +
  #      ggtitle("(d)") + # Assign panel number/header
  #      theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
   #           plot.title = element_text(face="bold")) + # This makes panel header bold 
      # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
        theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
)
ggsave(p9, filename="p9_Boxplot_Hb_by_Locality.pdf", height=7, width=9, units="in")
# As we'd expect, lots of locality differences; notably, Algarrobo (point 4) is super low! 

# Is there some sort of latitude effect????

# Oreotrochilus chimborazo color palette: 
# Deep purple: #6E1EDC
# Aqua gorget: #1CD2D8; another color: #0DAFA9
# pollen forehead: #D49A76
# gray patch brown: #665E54; another version #726A61; another version #8C847B
# deep green: #344D3F
# chuquiragua buds: #E2833F
# chuquiragua leaves
# snowy breast: #D1D1D9
scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87


# INTERACTION PLOT -GENERALIST SPECIALIST HB v. ELEV
(p10 <- ggplot(final, aes(x=elev, y=hb, colour=po2_class, linetype=po2_class)) +
# p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(data=final, aes(y=hb), size=2, alpha=0.8) +
    stat_smooth(data=final, aes(y=hb, group=po2_class), size=1.2, method="lm") +
    geom_jitter(width=0.8, height=0.8, alpha=0.8, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y="[Hb] g/dl", x="Elevation (m)") +
    #scale_color_manual(values=c(neither="black", generalist="burlywood2", specialist="#853FCD")) + #336B87
    scale_color_manual(values=c(neither="#8C847B", generalist="#1CD2D8", specialist="#6E1EDC")) + #336B87
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
   # ggtitle("(a)") + # Assign panel number/header
  #  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
  #        plot.title = element_text(face="bold")) + # This makes panel header bold 
    # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p10, filename="p10_Hb_vs_Elev_Scatterplot_PO2GeneralistSpecialist.pdf", height=7, width=9, units="in") 



# Now subset by just generalist and specialist and see what this looks like 
genspec <- final[which(final$po2_class == "generalist" | final$po2_class=="specialist"),]

# INTERACTION PLOT -GENERALIST SPECIALIST ONLY SUBSET HB v. ELEV
(p11 <- ggplot(genspec, aes(x=elev, y=hb, colour=po2_class, linetype=po2_class)) +
# p <- p + geom_boxplot(alpha = 0.5, outlier.size=0.1) # Add this back if categorical x-axis variable 
    geom_point(data=genspec, aes(y=hb), size=2, alpha=0.8) +
    stat_smooth(data=genspec, aes(y=hb, group=po2_class), size=1.2, method="lm") +
    geom_jitter(width=0.5, height=0.5, alpha=0.5, size=2) + # jitter size should match geom_point size or it looks weird
    labs(y="[Hb] g/dl", x="Elevation (m)") +
    scale_color_manual(values=c(generalist="#ffb20a", specialist="#853FCD")) + #336B87 (dark)
    theme_classic() +
    theme(legend.position = "right") + # Note: orange = spring, blue = winter 
   # ggtitle("(a)") + # Assign panel number/header
  #  theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
  #        plot.title = element_text(face="bold")) + # This makes panel header bold 
    # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
    theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10), axis.title=element_text(size=14))
)
ggsave(p11, filename="p11_Hb_vs_Elev_Scatterplot_PO2GeneralistSpecialistONLY.pdf", height=7, width=9, units="in") 

# light blue: #90AFC5; R 144, G 175, B 197
# dark teal: #336B87; R51, G107, B135
# purple orchid: #853FCD; R133, G63, B205


# Elev and Hb, faceted by species 
(p2 <- ggplot(final, aes(x=elev, y=hb)) + 
       facet_wrap(~species) +   # ok, way too many species right now to see any sort of clear pattern
       geom_point() +
       geom_smooth(method = "lm", colour="darkturquoise", size=1.2) + 
       theme_classic() + 
       labs(x="Elevation (m)", y="[Hb] g/dl") +
       # theme(axis.text.x = element_text(hjust = 1)) +
       theme(legend.position = "none") +
      # ggtitle("(a)") + # Assign panel number/header; this will be (a) because first in series of 3
      # theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
      #       plot.title = element_text(face="bold")) + # This makes panel header bold 
      #     # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
       theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=12), axis.title=element_text(size=14))
        )
#ggsave(p2, filename="m1.p1_TbMin_Vs_TaMin_Scatterplot_Simple.pdf", height=7, width=9, units="in")

# Boxplot of Hb v. sex w/ elevation
boxplot (hb ~ sex, data=final, xlab="Elevation (m)", ylab="[Hb] (g/dl)", boxwex=0.3, ylim=c(14,26))
stripchart(hb ~ sex, vertical = TRUE, data=full, method = "jitter", add = TRUE, pch = 20, col=c("gold", "purple", "olivedrab3"))
  # rgb(0.85,0.6,0,0.56),(rgb(0,0,0.95,0.4))))
# anovaHWIm <- aov(hb ~ sex, data=psub) 
# summary(anovaHWIm)
# text(1.5,24, "*", cex=3, col=("red"))

# I need to clean up the sex variable mega 
# There's also something weird about the way sexes are appearing: olive drab looks like female? (only a few unknown points)

# Hb v. elev
library(ggplot2)
p1 <- ggplot(final, aes(x=elev, y=hb))
p1 <- p1 + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p1 <- p1 + stat_smooth(method = lm)
p1 <- p1 + theme_bw()
p1 <- p1 + labs(title = "[Hb] v. elevation")
# p1 <- p1 + facet_wrap(. ~ country)
print(p1)

# Hct v. elev
library(ggplot2)
p1 <- ggplot(final, aes(x=elev, y=hct))
p1 <- p1 + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p1 <- p1 + stat_smooth(method = lm)
p1 <- p1 + theme_bw()
p1 <- p1 + labs(title = "Hct v. elevation")
# p1 <- p1 + facet_wrap(. ~ country)
print(p1)

# MCV v. elev
library(ggplot2)
p1 <- ggplot(final, aes(x=elev, y=mcv))
p1 <- p1 + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p1 <- p1 + stat_smooth(method = lm)
p1 <- p1 + theme_bw()
p1 <- p1 + labs(title = "MCV v. elevation")
print(p1)

# MCH v. elev
library(ggplot2)
p1 <- ggplot(final, aes(x=elev, y=mch))
p1 <- p1 + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p1 <- p1 + stat_smooth(method = lm)
p1 <- p1 + theme_bw()
p1 <- p1 + labs(title = "MCH v. elevation")
print(p1)


# MCHC v. elev
library(ggplot2)
p1 <- ggplot(final, aes(x=elev, y=mchc))
p1 <- p1 + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p1 <- p1 + stat_smooth(method = lm)
p1 <- p1 + theme_bw()
p1 <- p1 + labs(title = "MCHC v. elevation")
print(p1)


# Hb v. elev_position
library(ggplot2)
p <- ggplot(final, aes(x=elev_position, y=hb))
p <- p + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p <- p + stat_smooth(method = lm)
p <- p + theme_bw()
p <- p + labs(title = "[Hb] v. elevation_position")
# p1 <- p1 + facet_wrap(. ~ country)
print(p)

# Hct v. elev_position
library(ggplot2)
p <- ggplot(final, aes(x=elev_position, y=hb))
p <- p + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4)
p <- p + stat_smooth(method = lm)
p <- p + theme_bw()
p <- p + labs(title = "Hct v. elevation_position")
# p1 <- p1 + facet_wrap(. ~ country)
print(p)

```



# Plot for Chris Brown Bag: 
If you want to plot this, read in final.hct from modeling file
```{r}
# Scatterplot: Hct v. elev
(bbag_plot <- ggplot(final.hct, aes(x=elev_position, y=hct, colour=Clade)) + # Add shape=Clade here to get shapes to work properly
  geom_point(data=final.hct, size=1.5, alpha=0.8) +
  scale_colour_viridis(discrete=TRUE) + 
  #scale_color_manual(values = peru_chimborazo) +
 # geom_jitter(width=1.5, height=1.5, alpha=0.8, size=1.5) + # jitter to fill plot space 
  geom_smooth(method = "lm", colour="firebrick3", size=1.4) + # line used to be dark turquoise 
  theme_classic() + 
  labs(x="Elevation Position", y="Hct (%)") +
       # theme(axis.text.x = element_text(hjust = 1)) +
  theme(legend.position = "none") +
  # ggtitle("(b)") + # Assign panel number/header; this will be (a) because first in series of 3
  # theme(plot.title.position = "plot", # parameter "plot" specifies that you want "title" flush with y-axis
  #        plot.title = element_text(face="bold")) + # This makes panel header bold 
  #      # This is good for labeling figure panels! Avoids having to manually toy w/ hjust and vjust
  theme(axis.text.y=element_text(size=12), axis.text.x=element_text(size=12), axis.title=element_text(size=16))
 )
ggsave(bbag_plot, filename="bbag_plot_ScatterplotBasic_ElevPosition_by_Hct.pdf", height=3, width=4, units="in")
```



----- CUT -------

# 3D plots, SO cool 
but ultimately useless 
```{r}
library(plotly)
# holy fuck this is so cool

# 3d plot of elev, mcv, trbc
fig <- plot_ly(full1, x = ~elev, y = ~mcv, z = ~trbc, color = ~species)
fig <- fig %>% add_markers()
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Elevation (m)'),
                     yaxis = list(title = 'MCV'),
                     zaxis = list(title = 'RBC')))
fig

# 3d plot of hb with elev and mass
fig1 <- plot_ly(full1, x = ~elev, y = ~hb, z = ~mass, color = ~species)
fig1 <- fig %>% add_markers()
fig1 <- fig %>% layout(scene = list(xaxis = list(title = 'Elevation (m)'),
                     yaxis = list(title = '[Hb]'),
                     zaxis = list(title = 'Mass (g)')))
fig1
```






# Print environment for reproducibility
```{r}
sessionInfo() # List of packages and versions in use 
```

###########

## END 
